---
title: "fraud-detection-vignette"
author: "JP"
date: "29 July 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

Re-creation and expansion of [this blog](https://blog.codecentric.de/en/2017/09/data-science-fraud-detection/) 

Get data from [kaggle-data](https://www.kaggle.com/ntnu-testimon/paysim1).
Save in to the `\data` folder in the root of the repo and unzip. Re-name to 
`fd_data.csv`.

Get functions
```{r}
# LINE FOR INSTALLING FROM GITHUB HERE
library(JP.fraudDetection)

# remove once dependencies setup
library(data.table)
library(ggplot2)
library(ggridges)
```


## Data

```{r}
# load data
location = file.path(getwd(), "/data")
file = "fd_data.csv"
dt <- fread(file.path(location, file),
            colClasses = list(character=c("type","nameOrig","nameDest"),
                              numeric=c("amount","oldbalanceOrg","newbalanceOrig",
                                        "oldbalanceDest","newbalanceDest"),
                              integer=c("step","isFlaggedFraud"),
                              factor=c("isFraud"))
)
head(dt)
dim(dt)
# imbalance in classes: fraud is uncommon
dt[, .(.N, prop = .N/nrow(dt)), by = isFraud]
```


## Plots
sample the data for plotting, to save memory and speed
```{r}
p = 0.2
set.seed(1)
sample_dt <- rbindlist(list(dt[isFraud == 1,],
                            dt[isFraud == 0,][sample(1:.N, size = floor(.N*p)),])) 
sample_dt[, logAmount := log(amount + 1)]
```

Can see the amount is higher usually in Fraudulent cases
```{r}
pl <- ggplot(sample_dt, aes(x = logAmount, y = isFraud)) + 
  geom_density_ridges(fill = "goldenrod1") + theme_minimal()
pl
```

Only affects Transfer and Cash_out types of purchases
```{r}
pl <- ggplot(sample_dt, aes(x = logAmount, y = type, fill = isFraud)) + 
  geom_density_ridges(alpha = .6) + theme_minimal()
pl
prop_types <- dt[type %in% c("TRANSFER", "CASH_OUT"), .N/nrow(dt)]
```

These make up `r round(prop_types*100, 1)`% of the data 